---
title: "Project 3: MyStat302Package"
author: "Chaos Gao, Sherry Sheng"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MyStat302Package Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# Loading all the packages that are needed for this tutorial
library(MyStat302Package)
library(ggplot2)
library(dplyr)
library(class)
library(randomForest)
```

## Brief Introduction
For this MyStat302Package, you can install it by
```{r, eval=FALSE}
install.packages("MyStat302Package")
library(MyStat302Package)
```
or
```{r, eval=FALSE}
devtools::install_github("Chaos-Gao/MyStat302Package")
library(MyStat302Package)
```
In this package, it includes six functions, which are my_pow, f_to_c, my_t.test, my_lm, my_knn_cv, and my_rf_cv. Below contains four tutorials for my_t.test, my_lm, my_knn_cv, and my_rf_cv respectively.

## my_t.test Tutorial

```{r}
# A two tail t-test, with 60 as null hypothesis mean value, and alternative 
# hypothesis is mu not equal to 60
lifeExp_data <- my_gapminder$lifeExp
my_t.test(lifeExp_data, mu = 60)
```

* For the two side T-test, we have p value 0.09322, which is bigger than 0.05. Thus, we don't have sufficient evidence to reject the null hypothesis and can't support our alternative hypothesis.

```{r}
# A left tail t-test, with 60 as null hypothesis mean value, and alternative 
# hypothesis is mu smaller than 60
my_t.test(lifeExp_data, alternative = "less", mu = 60)
```

* For the left tailed T-test, we have p value 0.04661, which is smaller than 0.05. Thus, we have sufficient evidence to reject the null hypothesis and support our alternative hypothesis, which is mu < 60.

```{r}
# A right tail t-test, with 60 as null hypothesis mean value, and alternative 
# hypothesis is mu greater than 60
my_t.test(lifeExp_data, alternative = "greater", mu = 60)
```

* For the right tailed T-test, we have p value 0.95339, which is bigger than 0.05. Thus, we don't have sufficient evidence to reject the null hypothesis and can't support our alternative hypothesis, which is mu > 60

## my_lm Tutorial

```{r}
# Operate the linear regression
linear_pred <- my_lm(lifeExp ~ continent + gdpPercap, data = my_gapminder)
linear_pred
```
* The gdpPercap coefficient is 4.452704e-04. Holding other things constant, one unit increase in gdpPercap leads to 4.452704e-04 unit increase in lifeExp. 
* The null hypothesis is the coefficient of gdpPercap equals to zero. The alternative is the coefficient of gdpPercap does not equal zero. 
* The p value generated from the function is smaller than 0.05. Thus the result is statistically significant. Therefore, we reject the null hypothesis and conclude that there is a relationship between gdpPercap and lifeExp.

```{r, fig.width=5}
# Extract the coefficients
my_estimates <- linear_pred[, 1]
# Extract X variable
my_X <- model.matrix(lifeExp ~ continent + gdpPercap, my_gapminder)
# Calculate the fitted data
lifeExp_fit <- my_X %*% my_estimates

# Drew the Actual vs. Fitted plot
my_df <- data.frame(actual = my_gapminder$lifeExp, fitted = lifeExp_fit, 
                    continent = my_gapminder$continent)
ggplot(my_df, aes(x = fitted, y = actual, color = continent)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, col = "red", lty = 2) + 
  theme_bw(base_size = 15) +
  labs(x = "Fitted values", y = "Actual values", title = "Actual vs. Fitted") +
  theme(plot.title = element_text(hjust = 0.5))
```

* The fitted line cut through the middle value of the data points from each state. The points that lie above y = x have the actual value bigger than the fit value, and the points that lie below y = x have the fit value bigger than the actual value. However, many other random things also influence lifeExp, so that there are many outliers on the graph. So we conclude that the model is not fit well.

## my_knn_cv Tutorial

```{r}
# Remove NA in Data
penguins_data <- na.omit(my_penguins)
# Get the train and cl parameter from the data
train <- penguins_data[, 3:6]
cl <- penguins_data$species

# The list record the prediction
pred_class_list <- list()
# The vector record the test error
cv_err_vec <- vector()
# Do the knn prediction from 1 nearest neighbor to 10 nearest neighbors
for (j in 1:10) {
  temp <- my_knn_cv(train, cl, j, 5)
  pred_class_list[[j]] <- temp[[1]]
  cv_err_vec[j] <- temp[[2]]
}

# Calculate the training error
train_err_vec <- vector()
for (i in 1:10) {
  train_err_vec[i] <- sum(pred_class_list[[i]] != penguins_data$species) / 
    length(pred_class_list[[i]])
}

# Construct a table showing both the training and testing error
compare_err <- cbind.data.frame(cv_err_vec, train_err_vec)
compare_err
```
* Based on both the training misclassification rates and the CV misclassification rates, we would choose the 1 nearest neighbor cross-validation model.
* Also, in practice, we would choose 1 nearest neighbor cross-validation model, because it has the least the training misclassification rates and the CV misclassification rates.
* To apply the k-fold cross-validation, it first split data into k parts (folds). Then use all but 1 fold as your training data and fit the model and use the remaining fold for your test data and make predictions. Switch which fold is your test data and repeat previous steps until all folds have been test data (k times).Lastly, compute squared error.
* Cross-validation is important because it generally results in a less biased or less optimistic estimate of the model skill than all other methods.

## my_rf_cv Tutorial

```{r}
# Do the CV test with k = 2, 5, 8, and for each k, run 30 times
fold_vec <- c(2, 5, 8)
cv_err_mat <- matrix(nrow = 30, ncol = length(fold_vec))
for (i in 1:length(fold_vec)) {
  for (j in 1:30) {
    cv_err_mat[j, i] <- my_rf_cv(fold_vec[i])
  }
}
```

```{r, fig.width=5}
# Drew the three boxplots
cv_err_data <- as.data.frame(cv_err_mat)
colnames(cv_err_data) <- c("Fold2", "Fold5", "Fold8")
ggplot(data = cv_err_data) + 
  geom_boxplot(aes(x = "k = 2", y = Fold2)) +
  geom_boxplot(aes(x = "k = 5", y = Fold5)) +
  geom_boxplot(aes(x = "k = 8", y = Fold8)) +
  theme_bw(base_size = 10) +
  labs(title = "The MSE Distribution for k-folds Cross Validation",
       x = "Number of Folds",
       y = "CV Estimated MSE ") +
  theme(plot.title =
        element_text(hjust = 0.5))
```

```{r}
# Calculate the mean and sd of the MSE
mean_err_vec <- colMeans(cv_err_mat)
sd_err_vec <- vector()
for (i in 1:3) {
  sd_err_vec[i] <- sd(cv_err_mat[, i])
}

# Show the mean and sd in a table
cv_mean_sd <- cbind.data.frame(mean_err_vec, sd_err_vec)
cv_mean_sd
```
* According to the table and the box plot, the greater the value of k the smaller the mean error and the smaller the error's standard deviations. Since the more groups we split (bigger k) when doing CV, the more data are used to train our model. Therefore, our model would be more unbiased with smaller average test error. The standard deviation is also smaller when k increased. When k = 2, we only train our model with half of the sample data each time, which can lead to huge error difference since we split the data randomly. When k = 8, we might have our ideal trade-off bewteen bias and variance- the model we have each time is correlated moderatly and is trained with sufficient data, so the variance becomes lower. 

